### 2019-07-23 paper/intro.tex
small changes

\subsection{Gyrochronology}
The rotation periods of FGKM dwarfs increase with the square-root of time, and
this phenomenon was first observed several decades ago \citep{skumanich1972}.
This characteristic of main-sequence stars allows them to be dated via their
rotation periods, a practice known as gyrochronology.
This is convenient, since the ages of main-sequence stars are extremely
difficult, and at times impossible, to measure via the traditional age-dating
method of isochrone placement.
It has also been known for a long time that stars of the same age but
different masses have different rotation periods \racomment{(citation)}.
However, an underlying assumption behind many of the empirical gyrochronology
relations \citep[\eg][]{barnes2003, barnes2007, mamajek2008, meibom2011,
angus2015, angus2019} is that the relationships between rotation period and
color, and rotation period and age are {\it separable}.
In other words, the period-color relation is the same at all ages and the
period-age relation is the same at all colors.
It was recently shown that old stars appear to be too rapidly rotating for
their age, \citep{angus2015, vansaders2016, vansaders2018} and that a
mass-dependent modification to the classical \citet{skumanich1972} spin-down
law was needed in order to fit the data \citep{vansaders2016, vansaders2018}.
In addition, an even more recent analysis of middle-aged open clusters provide
further hints that the spin-down rate is mass-dependent \citep{curtis2019}.
Unfortunately, without rotation periods for members of old open clusters, or
any large group of old stars with precisely measured ages, it is extremely
difficult to calibrate the rotation-color relations at old ages.
For this reason, we turned to a population-based stellar age proxy: velocity
dispersion in order to investigate the period-temperature relations at old
ages.

>>>>

\subsection{Gyrochronology}
The rotation periods of FGKM dwarfs increase with the square-root of time, and
this phenomenon was first observed several decades ago \citep{skumanich1972}.
This characteristic of main-sequence stars allows them to be dated via their
rotation periods, a practice known as gyrochronology.
This is convenient, since the ages of main-sequence stars are extremely
difficult, and at times impossible, to measure via the traditional age-dating
method of isochrone placement.
It has also been known for a long time that stars of the same age but
different masses have different rotation periods \racomment{(citation)}.
However, an underlying assumption behind many of the empirical gyrochronology
relations \citep[\eg][]{barnes2003, barnes2007, mamajek2008, meibom2011,
angus2015, angus2019} is that the relationships between rotation period and
color, and rotation period and age are {\it separable}.
In other words, the period-color relation is the same at all ages and the
period-age relation is the same at all colors.
It was recently shown that old stars appear to be too rapidly rotating for
their age, \citep{angus2015, vansaders2016, vansaders2018} and that a
mass-dependent modification to the classical \citet{skumanich1972} spin-down
law was needed in order to fit the data \citep{vansaders2016, vansaders2018}.
In addition, an even more recent analysis of middle-aged open clusters
provides further hints that the spin-down rate is mass-dependent
\citep{curtis2019}.
Rotation period measurements of the 1.1 Gyr open cluster NGC 6811 reveal a
flattened relationship between rotation period and color, where the G dwarfs
are rotating at the same rate as the K dwarfs.

Unfortunately, without rotation periods for members of old open clusters, or
any large group of old stars with precisely measured ages, it is extremely
difficult to calibrate the rotation-color relations at old ages.
For this reason, we turned to a population-based stellar age proxy: velocity
dispersion in order to investigate the period-temperature relations at old
ages.

### 2019-07-23 paper/method.tex
small changes

\begin{figure}
  \caption{
A color magnitude diagram in SDSS g and 2MASS K bands showing the
\citet{mcquillan2014} sample, colored by their rotation periods.
We excluded photometric binaries and subgiants from our analysis by removing
stars above the two dashed lines.
The rotation periods of binaries and subgiants do not follow a Skumanich-like
braking law and were removed from our sample.
% since we rely on a simple scaling
% between rotation period and age to make the argument that the rotation period
% is not caused by incorrect period measurements.
}
  \centering
    \includegraphics[width=1\textwidth]{CMD_cuts}
\label{fig:CMD_cuts}
\end{figure}

In order to explore the age of this stellar population from a rotation-period
standpoint, it was first necessary to remove visual binaries and subgiants
from the sample.
The rotation-period evolution of these two types of stars is generally
different to that of single stars which more usually follow a Skumanich-like
spin-down law.
Tidal and magnetic interactions between the two components of a binary system
can influence the rotation periods of both stars, and the expanding envelopes
of subgiants drive rapid spin-down through conservation of angular momentum.
We removed visual binaries from the sample by fitting a 6th-order polynomial
to the main sequence and raising it by subtracting 0.15 magnitudes to
approximate the line between the single-star and binary main sequences.
This polynomial is shown in figure \ref{fig:CMD_cuts}.
\racomment{Why did we do this in g-K?}
In order to remove subgiants, we removed stars brighter than 6th magnitude in
the g-band from the sample.

>>>>

\begin{figure}
  \caption{
A color magnitude diagram in SDSS g and 2MASS K bands showing the
\citet{mcquillan2014} sample, colored by their rotation periods.
We excluded photometric binaries and subgiants from our analysis by removing
stars above the two dashed lines.
The rotation periods of binaries and subgiants do not follow a Skumanich-like
braking law and were removed from our sample.
The rotation period gradient across the main sequence is visible by eye in
    this figure: young, rapidly rotating stars are located below the old,
    slowly rotating stars.
% since we rely on a simple scaling
% between rotation period and age to make the argument that the rotation period
% is not caused by incorrect period measurements.
}
  \centering
    \includegraphics[width=1\textwidth]{CMD_cuts}
\label{fig:CMD_cuts}
\end{figure}

In order to explore the age of this stellar population from a rotation-period
standpoint, it was first necessary to remove visual binaries and subgiants
from the sample.
The rotation-period evolution of these two types of stars is generally
different to that of single stars which more usually follow a Skumanich-like
spin-down law.
Tidal and magnetic interactions between the two components of a binary system
can influence the rotation periods of both stars, and the expanding envelopes
of subgiants drive rapid spin-down through conservation of angular momentum.
We removed visual binaries and subgiants from the sample by applying cuts to
the color-magnitude diagram (CMD).
We fit a 6th-order polynomial to the main sequence and raised it by
subtracting 0.15 magnitudes to approximate the line between the single-star
and binary main sequences.
This polynomial is shown in figure \ref{fig:CMD_cuts}.
We removed all stars above this line from the sample.
\racomment{Why did we do this in g-K?}
In order to remove subgiants, we removed stars brighter than 6th magnitude in
the g-band from the sample.
